S-Aligner: Ultra-scalable read mapping on Sunway Taihu Light

### 2017-04-05 text/design.tex
move content to text file.

\documentclass[conference]{IEEEtran}

\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage{cases}
\usepackage{listings}
\lstset{language=C++,
    basicstyle=\ttfamily \footnotesize,
    keywordstyle=\color{black}\ttfamily,
    stringstyle=\color{black}\ttfamily,
    commentstyle=\color{gray}\ttfamily,
    morecomment=[s][\color{blue}]{/**}{**/},
    showstringspaces=false,
    showtabs=false,
    captionpos=b,
    numbers=left
}

\hyphenation{cluster computational-biology short-read-mapping}

\begin{document}

\title{S-Aligner: Ultra-scalable read mapping on Sunway Taihu Light}

\author{\IEEEauthorblockN{Xiaohui Duan, Kai Xu, Yuandong Chan}
\IEEEauthorblockA{%College of Computer Science and Technology \\
Shandong University\\
Jinan, China\\
Email: \{sunrise.duan, xu\_kai, yuandong.c\}@mail.sdu.edu.cn}
\and
\IEEEauthorblockN{Christian Hundt, Bertil Schmidt}
\IEEEauthorblockA{Johannes Gutenberg University\\
Mainz, Germany\\
Email: \{hundt, bertil.schmidt\}@uni-mainz.de}
\and
\IEEEauthorblockN{Pavan Balaji}
\IEEEauthorblockA{Argonne National Laboratory\\
Argonne, Illinois, USA\\
Email: balaji@anl.gov}
\and
\IEEEauthorblockN{Weiguo Liu}
\IEEEauthorblockA{Shandong University\\
Jinan, China\\
Email: weiguo.liu@sdu.edu.cn\\
}
}

\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The availability and amount of sequenced genomes have been rapidly growing in recent years due to the adoption of next-generation sequencing (NGS) technologies that allow for high-throughput short-read generation at highly competitive cost. Since this trend is expected to continue in the foreseeable future, the design and implementation of efficient and scalable NGS bioinformatics algorithms are important to research and industrial applications.  In this paper, we introduce S-Aligner---a highly scalable read mapper designed for the \emph{Sunway Taihu Light} supercomputer and its fourth-generation ShenWei many-core architecture (SW26010). S-Aligner employs a combination of optimization techniques to overcome both the memory-bound and the compute-bound bottlenecks in the read mapping algorithm. In order to make full use of the compute power of Sunway Taihu Light, our design employs three levels of parallelism: (1) inter-node parallelism using MPI based on a task-grid pattern, (2) intra-node parallelism using multi-threading and asynchronous data transfer to fully utilize all 260 cores of the SW26010 many-core processor, and (3) vectorization to exploit the available 256-bit SIMD vector registers.  Moreover, we have employed asynchronous access patterns and data sharing strategies during file I/O to overcome bandwidth limitations of the network file system. Our performance evaluation demonstrates that S-Aligner scales almost linearly with approximately 95\% efficiency for up to 13,312 nodes (concurrently harnessing more than 3 million compute cores). Furthermore, our implementation on a single node outperforms the established RazerS3 mapper running on a platform with eight Intel Xeon E7-8860v3 CPUs while achieving highly competitive alignment accuracy.
\end{abstract}

\IEEEpeerreviewmaketitle

\section{Introduction}
\label{Introduction}

{\em Next-generation sequencing} (NGS) technologies have revolutionized the field of computational biology because of their massive throughput at highly competitive cost. Nowadays, NGS data is processed in almost every field of bioinformatics with important applications including personalized cancer treatment and precision medicine.
Modern NGS instruments such as Illumina sequencers \cite{reviewngs} allow for the recording of billions of short DNA fragments per day. The produced \emph{short reads} are usually only a few hundred base pairs (bps) in length compared with complete genomes of typical mammals covering billions of nucleotides. In order to further analyze the set of delocalized reads, they are typically mapped to a given reference genome by computing base-pair level alignments to determine their original positions. This exhaustive probing of possible alignments is a compute-heavy task and thus justifies the need for massively parallel and efficient computation patterns. Moreover, read alignment (or mapping) is crucial for many DNA sequence analysis pipelines such as genotyping, the discovery of single nucleotide polymorphism, or personal genomics.

The na\"ive computation of optimal local alignments between each read and the reference genome is considered prohibitive because of the demanding time complexity of the Smith-Waterman algorithm \cite{sw} that is proportional to the product of the sequences' lengths. Thus, methods based on a {\em seed-and-extend} approach are typically employed to dramatically reduce the number of alignment positions by considering only those regions in the reference genome which contain an exact occurrence of a substring of a read. Among existing read mappers one can find two predominant operation modes: {\em any-best} and {\em all}. Bowtie2 \cite{bowtie2}, BWA \cite{bwa}, GEM \cite{gem}, and CUSHAW \cite{cushaw} run in {\em any-best} mode, which determines only one or a few best alignments (also called {\em intervals}) of each read to the reference genome. Other mappers such as RazerS3 \cite{razers3}, Hobbes2 \cite{hobbes2}, BitMapper \cite{bitmapper}, and mrFAST \cite{mrfast} operate in {\em all} mode, which reports all intervals satisfying a certain constraint, for example, compliance with a predefined edit distance threshold.  Because of their extensive nature all-mappers demand far more computational resources than their any-best counterparts. Nevertheless, when studying gene innovation or phenotypic variation the complete list of alignments is crucial for a thorough analysis.

Consequently, we pursue the design of an all-mapper in order to support a wide range of biomedical applications. Associated runtimes can be dramatically reduced if we target clusters with a large number of compute nodes. In the recent past, the prevalent trend in supercomputing has favored heterogeneous nodes featuring attached coprocessors, such as GPUs or Xeon Phi architectures, because of their vast compute capabilities and low energy footprint. A similar architecture has recently been introduced with the \emph{Sunway Taihu Light} supercomputer, which is based on the neo-heterogeneous SW26010 many-core processor. This cluster consists of 40,960 nodes. Each node provides 260 compute cores organized in four {\em compute groups} (CGs). A total of 10,649,600 cores provides Linpack performance of up to 93 PFlop/s, rendering it the world's fastest supercomputer in the Top500 list of November 2016.

In this paper, we introduce S-Aligner---a highly scalable read mapper targeting the Sunway Taihu Light supercomputer and its SW26010 architecture. To the best of our knowledge, we are the first to investigate the applicability and scalability of read mapping on this architecture. Our approach relies on four major contributions. First, we provide an efficient implementation of Myers' bit-parallel string matching algorithm that harnesses the full parallelization and vectorization potential of the SW26010 many-core processor. Our new implementation is three orders of magnitude faster than a na\"ive single-threaded Myers' version. Second, we exploit fast local device memory via DMA intrinsics for intra-CG communication, which results in a 22$\times$ speedup compared with the non-DMA variant. Third, a highly scalable inter-CG communication scheme is proposed. This results in a parallelization efficiency of over 95\% when using 53,248 CGs. Fourth, we employ asynchronous file loading and data-sharing strategies to effectively hide the latency of the network file system. We further expect that the presented techniques can guide the design and implementation of similar types of applications on the neo-heterogeneous many-core cluster architecture of Sunway Taihu Light.

The rest of this paper is organized as follows: Section~\ref{Related Work} describes required background about the Sunway Taihu Light architecture, related work, and the seed-and-extend approach to read mapping. Details of the design and implementation of S-Aligner are presented in Section~\ref{Implementation}. Performance evaluations in terms of speed, scalability, and alignment quality are carried out in Section~\ref{Evaluation}. Section~\ref{Conclusion} summarizes our conclusions.

\section{Background}
\label{Related Work}

\subsection{Sunway Taihu Light Architecture}
\label{Sunway Taihu Light}

The Sunway Taihu Light has been manufactured by the National Research Center of Parallel Computer Engineering \& Technology (NRCPC) of the People's Republic of China and is located at the Wuxi Supercomputing Center. It provides a theoretical peak performance of 125 PFlop/s and an effective performance-to-power ratio of over 6 GFlop/s per watt. It consists of 40,960 nodes with 1.4 PB attached memory. The interconnection network provides roughly 70 TB/s bisection bandwidth and 12 GB/s point-to-point communication bandwidth.

\begin{figure}[!htb]
	\begin{center}
		\includegraphics[width=\linewidth]{SW26010}
		\caption{Architecture of the SW26010 processor, which is partitioned into four CGs.}
		\label{fig:SW26010}
	\end{center}
\end{figure}

Each node is equipped with a single SW26010 processor that is subdivided into four {\em compute groups} (CGs). A CG is the basic unit that can be addressed by the job scheduling system. Each CG consists of a single {\em master processor} (MP), 64 {\em slave processors} (SPs), and 8 GB of attached DDR3 shared memory. All nodes can access a shared network file system for loading and storing data. The operating system is a customized Linux flavor running on the MP.  Users may manually launch threads on the SPs in order to parallelize compute-heavy portions of their code. Both the MP and 64 SPs support ShenWei's RISC basic instruction set, which provides scalar and SIMD operations. Moreover, reduction primitives can be called exclusively on the MP, while the SPs exhibit specialized intrinsics for high-precision integer arithmetic. The 8 GB attached shared memory can be accessed by both the MP and the 64 SPs via a memory controller with a shared bandwidth of approximately 136 GB/s. The MP has its own cache (L1 and L2), and each SP can access 64 KB of fast {\em local device memory} (LDM). Data residing in shared memory can be written to LDM using DMA intrinsics and subsequently communicated via a broadcast to the LDM of other SPs. Figure~\ref{fig:SW26010} shows the described hardware layout.

ShenWei's SIMD instruction set features 256-bit wide vector types and a set of corresponding intrinsics provided by the \emph{ShenWei 5 Compiler Collection}.  The MPI library (Sunway MPI) is a derivative of the MPICH implementation, and is compliant with the MPI-3 standard.  Basic threading capabilities for the SPs are provided by the \emph{athread} library.

\subsection{Related Work}
We are not the first to parallelize read mapping on a compute cluster. Among the earliest tools are pBWA \cite{pbwa} and pMap \cite{pmap}. pBWA is an MPI implementation of Version 0.5.9 of the BWA aligner while pMap provides an MPI-based framework for the concurrent execution of popular aligners such as BWA \cite{bwa} and Bowtie \cite{bowtie}. Other implementations exploit MapReduce frameworks, for example, the Hadoop-based SEAL \cite{seal} and BigBWA \cite{bigbwa} tools or the Spark-based SparkBWA suite \cite{sparkBWA}, in order to guarantee fault tolerance. Unfortunately, all these approaches suffer from insufficient scalability when executed on hundreds of nodes. The metaframework parSRA \cite{parSRA} is designed for read mappers written in UPC++.  It combines dynamic scheduling and a virtual file system layer for read distribution to overcome the limitations imposed by pMap and MapReduce-based approaches. merAligner \cite{merAligner} and CUSHAW3-UPC++ \cite{cushaw3upc} are other PGAS-based short-read aligners written in UPC; both demonstrate good scalability. Although pMap and parSRA allow for the execution of single node GPU aligners such as nvBowtie \cite{nvBio}, none of the cited tools is explicitly designed for heterogeneous clusters consisting of thousands of many-core processors. Furthermore, all previous approaches target traditional hardware architectures.

To the best of our knowledge, S-Aligner is the first attempt to implement a fully scalable read mapper specifically designed to fit the characteristics of Sunway Taihu Light. Previous algorithms mapped onto this novel supercomputer have focused mainly on application domains outside bioinformatics, such as Earth system modeling, ocean surface wave modeling, atomistic simulation, and phase-field simulation \cite{sunway}.

\subsection{Seed-and-Extend Approach}
%There are some "all" mappers currently on traditional commercial platforms, RazerS3\cite{razers3} is one of the most famous "all" mapper on x86\_64 CPUs. Hobbes2\cite{hobbes2}, BitMapper \cite{bitmapper} and mrFAST\cite{mrfast} are also read mappers in "all" mode on x86\_64 platforms. BitMapper2 also supports CUDA-enabled GPUs.

Consider a set of reads ${\cal R}$, a reference genome $G$, and an error threshold $e$. The read mapping problem can be defined as follows: Find all substrings $g$ of $G$ that are within edit distance $e$ to some read $R \in {\cal R}$. We call such occurrences $g$ in $G$ {\em matches}.

This problem can be solved by a classical {\em dynamic programming} (DP) approach that calculates the semi-global alignment between each $R \in {\cal R}$ and $G$. Unfortunately, each alignment results in a time complexity proportional to the product of the sequences' lengths, which renders intractable the alignment of a large number of short reads to a reference genome a few billion letters long.  In order to address this problem, most state-of-the art solutions \cite{Reinert} employ the {\em seed-and-extend} approach, which follows a three-stage pipeline.

\begin{description}

\item[Stage 1: Filtration.] This stage identifies promising candidate regions (called {\em seeds}) for each read in $G$. A popular strategy in order to discard large regions of $G$ is based on the fact that if a read $R \in {\cal R}$ is divided into $e+1$ non-overlapping $q$-grams (substrings of length $q = \lfloor \lvert R\rvert\ /(e+1) \rfloor$), then (according to the pigeonhole principle) at least one of them occurs exactly in a match. Such occurrences can be identified quickly by looking them up in precomputed $q$-gram index data structure (also called the {\em reference genome index}), which stores all substrings of length $q$ of $G$.

\item[Stage 2. Verification.] This stage determines whether a seed can actually be extended to a full match within edit distance $e$. This requires the implementation of a {\em verification} algorithm in order to analyze the vicinity of each seed. Most mappers typically apply fast versions of DP-based algorithms for this step.

\item[Stage 3. Alignment.] This stage generates the base-pair level alignment information of a read and its verified intervals in $G$. 
%{\color{red} Note that some mappers just use a traceback of Myers delta matrix and some use Smith-waterman, so we shouldn't mention the algorithm}
\end{description}

Established all-mappers such as RazerS3 \cite{razers3} and mrFAST \cite{mrfast} follow this pipeline. Our profiling of RazerS3 and mrFAST using a typical benchmark data set including the human reference HG19 and 1 million simulated Illumina reads (with a read length of 100 bps) reveals that Stage 2 occupies 67\% and 93\% of the overall runtime, respectively. This shows that a fast implementation of Myers' bit-parallel algorithm for computing the edit distance between a read and an interval is a crucial component when designing efficient all-mappers.

Consider two strings $s$ and $s^\prime$ of length $n$ and $m$, respectively. Their {\em edit distance} is the minimum number of point mutations (i.e. insertions, deletions, or substitutions) required to transform $s$ into $s^\prime$. It can be determined by relaxing the cells of a cost matrix $C$ of size $(n+1) \times (m+1)$ according to the recurrence relation shown in Equation~\ref{update_scheme}, where $0 < i \le n$ , $0 < j \le m$ and the {\em characteristic function} $\chi (x \ne y)$ being 1 if $x \ne y$, and 0 otherwise. Initial conditions are set to $C[i,0] = i$, $C[0,j] = j$, and $C[0,0] = 0$.

\begin{align} \label{update_scheme} 
C[i,j]=\min \begin{cases} C[i-1,j-1]+\chi(s_{i-1} \neq s'_{j-1})\\
C[i-1,j]+1 \\
C[i,j-1]+1
\end{cases}
\end{align}

Myers proposed a bit-vector algorithm \cite{myers} that exploits bit parallelism by encoding the differences (deltas) between adjacent rows and columns in the cost matrix $C$ as defined in Equation~\ref{delta_update}, where $\varDelta h_{i,j}$,  $\varDelta v_{i,j}$, and  $\varDelta d_{i,j}$ are the discrete derivatives of $C[i, j]$ in the horizontal, vertical, and diagonal direction, respectively.

\begin{align} \label{delta_update}
\varDelta v_{i,j}&=C[i,j]-C[i-1,j]   &&\in \{0, \pm1\} \nonumber\\
\varDelta h_{i,j}&=C[i,j]-C[i,j-1]   &&\in \{0, \pm1\} \\
\varDelta d_{i,j}&=C[i,j]-C[i-1,j-1] &&\in \{0, +1\}   \nonumber
\end{align}

Note that the absolute values in Equation~\ref{delta_update} are either 0 or 1. Thus, we can encode the relatively small state space with the help of five vectors using one-hot encoding. Figure~\ref{BitDP} shows an example for the encoding of the vertical derivative $\varDelta v$ and its associated one-hot representations $V^+$ and $V^-$.  With this bit-vector representation the cell updates can be rewritten in terms of logical operations as shown in Equation~\ref{one-hot-myers}:

\begin{alignat}{100} \label{one-hot-myers}
H^-_{i, j} &= \chi(\varDelta h_{i, j} &&= -1)            &&= V^+_{i, j-1} &&\land D^0_{i, j} \nonumber\\
V^-_{i, j} &= \chi(\varDelta v_{i, j} &&= -1)            &&= H^+_{i-1, j} &&\land D^0_{i, j} \nonumber\\
H^+_{i, j} &= \chi(\varDelta h_{i, j} &&= +1)            &&= V^-_{i, j-1} &&\lor \lnot(V^+_{i, j-1} \,\lor D^0_{i, j}) \\
V^+_{i, j} &= \chi(\varDelta v_{i, j} &&= +1)            &&= H^-_{i-1, j} &&\lor \lnot(H^+_{i-1, j}   \lor D^0_{i, j}) \nonumber\\
D^0_{i,j}  &= \chi(\varDelta d_{i, j} &&= \phantom{+}0)  &&= V^-_{i, j-1} &&\lor H^-_{i-1, j} \lor \chi(s_{i-1} = s'_{j-1}) \nonumber 
\end{alignat}

\begin{figure}[!htb]
	\includegraphics[width=1\linewidth]{BitDP}
	\caption{Relations between the original matrix, $\varDelta v$, $V^+$, and $V^-$.  Here (a) shows the original score matrix; light-shaded cells indicate $\varDelta v_{i,j}=+1$ while heavily shaded cells correspond to $\varDelta v_{i,j}=-1$. The vertical derivative $\varDelta v$ shown in (b) is subsequently one-hot encoded by the bit vectors $V^+$ and $V^-$ in (c) and (d), respectively.}
	\label{BitDP}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=\linewidth]{GridNew}
	\caption{Our task-partitioning and file-loading strategy: (a) overall design; (b) and (c) detailed behavior of processes.}
	\label{TaskGrid}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=\linewidth]{AsyncRead}
	\caption{Asynchronous data loading strategy: (a) default case, and (b) case when there is no significantly shorter reference block.}
	\label{AsyncRead}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=1\linewidth]{FrmWk}
	\caption{Workflow of S-Aligner on a single CG: gray boxes correspond to tasks assigned to SPs while tasks in white boxes are executed on the MP.}
	\label{FrmWk}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=1\linewidth]{AsyncPara}
	\caption{Asynchronous filtration.}
	\label{AsyncPara}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=\linewidth]{SPParallel}
	\caption{Framework of our parallelization scheme within one CG: the MP performs filtration and output of alignment results while SPs verify identified candidate locations and generate base-pair level alignments. Gray boxes indicate the intervals verified in the second round of verification (assuming $slice\_num = 4$).}
	\label{SPParallel}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=0.9\linewidth]{MixPack}
	\caption{Bit-level encoding strategies for DNA sequences. For simplicity, we use a 4-bit word instead of 64-bit word for illustrating our combined AoSoA strategy (bottom).}
	\label{MixPack}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=1\linewidth]{AsyncTrans}
	\caption{Asynchronous data transfer to LDM.}
	\label{AsyncTrans}
	%\end{center}
\end{figure}

@@

\section{Performance Evaluation}
\label{Evaluation}
The performance of S-Aligner has been evaluated on the Sunway Taihu Light supercomputer. We use GRCh38\footnote{available at http://hgdownload.cse.ucsc.\\edu/downloads.html} as the human reference genome.  For the read input data sets we use either ERR013135\footnote{available at ftp://ftp.sra.ebi.ac.uk/vol1/fastq/ERR013/\\ERR013135} or reads simulated by Mason \cite{mason} or wgsim \cite{wgsim}.\footnote{Note that Mason usually generates reads of higher quality but does not support long reads.}  We first evaluate the performance in terms of runtime and mapping accuracy on a single node. Subsequently, we evaluate the scalability of our implementation by varying the number of utilized nodes up to 13,312.

\subsection{Single-Node Performance Analysis}

In this experiment, we used the first chromosome of GRCh38 as reference and 20K reads of length 200 bps generated by Mason as input data.

First, we evaluated the runtime performance of four different implementations of edit distance calculation using Myers' algorithm in one CG: (1) a na\"ive method using the MP only, (2) a na\"ive method with a non-vectorized SP parallelization, (3) a vectorized Myers' algorithm with the MP only, and (4) our vectorized Myers' algorithm with SP parallelization.

\begin{figure}[!htb]
	\begin{center}
		\includegraphics[width=1\linewidth]{VarVerCha}
		\caption{Comparison of the runtimes of four implementations of Myers' algorithm on a single CG.}
		\label{VarVerCha}
	\end{center}
\end{figure}

The result shown in Figure \ref{VarVerCha} demonstrates that making full use of SPs and bit-parallel SIMD vectorization gains dramatic speedups (118$\times$ and 63$\times$, respectively). Since the MP does not support high-precision (256-bit integer) extensions, we implemented a workaround for the high-precision addition operation. This makes it run slower on a single MP than on a single SP. Thus, using SPs can gain a superlinear speedup in this application. Furthermore, our bit-parallel SIMD implementation is able to update 256 cells simultaneously using 23 instructions (i.e., 0.09 instructions per cell) while the na\"ive implementation computes one cell using 6 instructions. This results in a theoretical speedup of $\sim$67$\times$, which explains the actually achieved speedup of 63$\times$.

Second, we evaluated the impact of using asynchronous filtration and the impact of using DMA intrinsics versus \texttt{memcpy}. The result is shown in Figure \ref{VarOptCha}.

\begin{figure}[!htb]
	\begin{center}
		\includegraphics[width=1\linewidth]{VarOptCha}
		\caption{Evaluation with and without I/O optimizations.}
		\label{VarOptCha}
	\end{center}
\end{figure}

The result demonstrates that DMA fetching is key for achieving high performance when using SPs, since a speedup of $\sim$22$\times$ is gained over \texttt{memcpy}. This can be attributed to two factors: (1) DMA transfers require fewer compute resources since they are performed by the memory controller, whereas \texttt{memcpy} uses an SP to write data to memory; (2) DMA fetching can be performed asynchronously, thus enabling the latency of fetching data to be hidden by computation. Furthermore, asynchronous filtration gains $\sim$15\% speedup.

In summary, the architecture of SW26010 differs significantly from a conventional x86\_64 CPU. Straightforward migration of code therefore typically results in an inefficient implementation on SW26010, and architecture-specific optimizations are required in order to achieve high performance. As a case study, we compiled and executed multithreaded BWA \cite{bwa} (one of the most well-known {\em any-best mappers}) on SW26010. It runs two times slower than S-Aligner, while finding significantly fewer locations, showing the importance of making changes according to this specific architecture.

\subsection{Comparison with RazerS3}
In this experiment, we use the first chromosome from GRCh38 as reference and various number of reads from ERR013135. The runtime of S-Aligner on a single SW26010 node is compared with RazerS3, a representative all-mapper. Since RazerS3 does not support ShenWei's architecture, we ran it on a machine with Intel processors using multithreading with default parameters (e.g., the accuracy parameter is set to 98\%). Measured runtimes are shown in Table \ref{SingleNode}.  One can see that S-Aligner executed on a single SW26010 (260 cores running at 1.45 GHz) outperforms RazerS3 running on eight Xeon E7-8860v3 CPUs (128 cores running at 3.2 GHz).

\begin{table}
	\begin{threeparttable}
		\caption{Runtime comparison between S-Aligner running on a single CG and RazerS3 running on a single node with eight Xeon E7-8860v3 CPUs.}
		\label{SingleNode}
		\begin{tabular}{@{\extracolsep{2pt}}rrrrr}
			\hline
			\multicolumn{1}{c}{Ref} &
			\multicolumn{2}{c}{Reads} &
			\multicolumn{2}{c}{Time(s)}\\
			\cline{2-3}
			\cline{4-5}
			\multicolumn{1}{c}{\#bps} &
			\multicolumn{1}{c}{Count} &
			\multicolumn{1}{c}{\#bps} &
			\multicolumn{1}{c}{RazerS3\tnote{\textdagger}} &
			\multicolumn{1}{c}{S-Aligner\tnote{\textdaggerdbl}}\\
			\hline
			%65M & 10M  & 108 &  52\tnote{\textdagger} & 47\tnote{\textdaggerdbl}\\
			%58M & 40M  & 108 &  1415 & 435\\
			116M & 40M & 108 &  939 & 405\\
			253M & 40M & 108 &  3,044 & 892\\
			%58M & 40M  & 200 &  898 & 1,002\\
			%116M & 40M & 200 &  744 & 1,013\\
			253M & 40M & 200 &  3,430 & 2,580\\
			\hline
		\end{tabular}
		\begin{tablenotes}
		\item[\textdagger] result is from a machine with eight Xeon E7-8860v3 (128 cores up to 3.20GHz).
		\item[\textdaggerdbl] result is from a node with a SW26010 (260 cores with frequency of 1.45GHz). 
		\end{tablenotes}
	\end{threeparttable}
\end{table}

Rabema \cite{rabema} (Read Alignment BEnchMark) is a well-defined read alignment benchmark that can evaluate the quality of read mappers in both {\em all} and {\em any-best} mode. We evaluated the accuracy of S-Aligner in all mode based on Rabema by using the result of RazerS3 with the accuracy parameter set to 100\% as gold standard. S-Aligner can find up to 99.8\% of the normalized intervals found by RazerS3 (100\% accuracy), which is higher than the accuracy of RazerS3 executed with default parameters (98\% accuracy). Detailed results are provided in Table~\ref{AccuEval}.

%As we can see, our algorithm has better performance than RazerS3, with an accuracy no worse than ordinary RazerS3.

\begin{table}
	\begin{threeparttable}
		\caption{Accuracy evaluation of S-Aligner for both real and simulated data.}
		\label{AccuEval}
		\begin{tabular}{@{\extracolsep{2pt}}rrrrrr}
			\hline
			\multicolumn{1}{c}{Chrom} &
			\multicolumn{3}{c}{Reads} &
			\multicolumn{1}{c}{\multirow{2}{*}{Interv. found}} \\
%			\multicolumn{1}{c}{Normalized}\\
			\cline{2-4}
			\multicolumn{1}{c}{Index} &
			\multicolumn{1}{c}{Origin} &
			\multicolumn{1}{c}{Count} &
			\multicolumn{1}{c}{\#bps} \\		
%			\multicolumn{1}{c}{intervals found}}\\
			\hline
			%65M & 10M  & 108 &  52\tnote{\textdagger} & 47\tnote{\textdaggerdbl}\\
			%58M & 40M  & 108 &  1415 & 435\\
			1\textsuperscript{st} & ERR013135 & 20,000 & 108 & 99.34\%\\
			1\textsuperscript{st} & Mason &  20,000 & 200 & 99.82\%\\
			\hline
		\end{tabular}
	\end{threeparttable}
\end{table}

\subsection{Scalability Analysis}

To evaluate weak and strong scalability, we measured the runtimes of S-Aligner using different numbers of nodes ranging from 13 to 13,312. The full GRCh38 assembly of the human genome was used as reference. Input reads were simulated by Mason with an error rate of 4\% and a length of 200 bps. Weak scalability was measured for numbers of nodes ranging from 13 to 3,328 by increasing the number of input reads from 16 million to 4 billion, correspondingly. Strong scalability was measured by increasing the number of nodes from 3,328 to 13,312 while keeping the number of input reads constant at 4 billion. We measured the performance in terms of {\em read base-pairs processed per second} (bpps) in total and normalized per node (see Table \ref{paraexp}). Figure \ref{VarProcCha} shows the achieved speedups and efficiencies. The results demonstrate weak scalability since the normalized node performance is almost constant for the number of nodes ranging from 13 to 3,328. Furthermore, the node performance only slightly decreases when increasing the number of nodes from 3,328 to 13,312 while keeping the amount of input reads constant at 4 billion, thus demonstrating strong scalability for sufficiently large input datasets.

\begin{table}[!htb]
	\begin{threeparttable}
		\caption{Performance and runtime evaluation of S-Aligner using different numbers of nodes.}
		\label{paraexp}
		\begin{tabular}{rrrrrr}
			\hline
			\multicolumn{1}{c}{\multirow{2}{*}{\#nodes}} &
			\multicolumn{2}{c}{Reads}  &
			\multicolumn{1}{c}{\multirow{2}{*}{Time(s)}} &
			\multicolumn{2}{c}{Performance (bpps)} \\
			\cline{2-3}
			\cline{5-6}
			\multicolumn{1}{c}{} &
			\multicolumn{1}{c}{Count} &
			\multicolumn{1}{c}{\#bps} & &
			\multicolumn{1}{c}{Total (M)} &
			\multicolumn{1}{c}{Node (K)}\\
			\hline
			13 & 16M & 200 & 1,315 & 2.43 & 185.43\\
			%			26 & 64 & 2664 & 4.9 & 187\\
			52 & 64M & 200 & 1,321 & 9.69 & 184.80\\
			%			104 & 64 & 2664 & 4.9 & 187\\
			208 & 256M & 200 & 1,321 & 38.76 & 184.66\\
			%			416 & 64 & 2664 & 4.9 & 187\\
			832 & 1,024M & 200 & 1,327 & 154.33 & 184.59\\
			%			1664 & 64 & 2664 & 4.9 & 187\\
			3,328 & 4,096M & 200 & 1,349 & 607.26 & 179.67\\
			13,312 & 4,096M & 200 & 344 & 2,381.40 & 178.90\\
			\hline
		\end{tabular}
		\begin{tablenotes}
			\item
			\#nodes is the number of nodes involved in computing.
			\item
			bpps is short for million bp processed per second.
			\item
			M / K indicates that column is presented in millions / thousands.
		\end{tablenotes}
	\end{threeparttable}
\end{table}

\begin{figure}[!htb]
	\begin{center}
		\includegraphics[width=1\linewidth]{VarProcCha}
		\caption{Speedup and efficiency of S-Aligner for different numbers of nodes.}
		\label{VarProcCha}
	\end{center}
\end{figure}

\section{Conclusion}
\label{Conclusion}

NGS data will continue to grow rapidly in the foreseeable future. Read mapping is a critical and compute-intensive step for a variety of NGS pipelines. While significant efforts have been devoted to optimizing this task, it is still a major bottleneck. In this paper, we have introduced S-Aligner---a highly scalable read mapper specifically designed to fit the characteristics of the Sunway Taihu Light supercomputer and its SW26010 architecture.
It scales almost linearly with over 95\% parallelization efficiency when distributed over more than 50,000 processes. As a result, S-Aligner can map $\approx$ 1.6 TB of read data to the whole human genome in just a few minutes. Moreover, when executed on a single node it can outperform the popular all-mapper RazerS3 executed on eight Xeon E7-8860 CPUs while achieving highly competitive mapping accuracy.

The design of S-Aligner teaches a number of lessons about properly designing parallelization and communication patterns in order to achieve both high performance and scalability on the new Sunway Taihu Light supercomputer. First, both multithreading and SIMD vectorization need to be employed to fully exploit the computational resources of the SW26010 processor. Second, within an SP the fast LDM must be used via DMA intrinsics in order to guarantee efficient intra-CG communication. Third, a scalable inter-CG communication scheme must be implemented in a way that efficiently hides the interconnection network bottleneck. Fourth, asynchronous file-loading and data-sharing strategies need to be implemented in order to effectively hide the latency of the network file system. The techniques presented in this paper could also be adapted to map similar applications exhibiting a pipeline of hashing and verification, such as large-scale approximate near duplicate object detection \cite{efficient} or de novo genome assembly \cite{swap} onto the heterogeneous many-core cluster architecture of Sunway Taihu Light.

\bibliographystyle{plain}
\bibliography{references.bib}

% that's all folks
\end{document}

>>>>

@@

\begin{figure}[!htb]
  \includegraphics[width=\linewidth]{figures/GridNew}
  \caption{Our task-partitioning and file-loading strategy: (a)
    overall design; (b) and (c) detailed behavior of processes.}
  \label{TaskGrid}
\end{figure}

@@

\begin{figure}[!htb]
  \includegraphics[width=\linewidth]{figures/AsyncRead}
  \caption{Asynchronous data loading strategy: (a) default case, and
    (b) case when there is no significantly shorter reference block.}
  \label{AsyncRead}
\end{figure}

@@

\begin{figure}[!htb]
  \includegraphics[width=1\linewidth]{figures/FrmWk}
  \caption{Workflow of S-Aligner on a single CG: gray boxes correspond
    to tasks assigned to SPs while tasks in white boxes are executed
    on the MP.}
  \label{FrmWk}
\end{figure}

@@

\begin{figure}[!htb]
  \includegraphics[width=1\linewidth]{figures/AsyncPara}
  \caption{Asynchronous filtration.}
  \label{AsyncPara}
\end{figure}

@@

\begin{figure}[!htb]
  \includegraphics[width=\linewidth]{figures/SPParallel}
  \caption{Framework of our parallelization scheme within one CG: the
    MP performs filtration and output of alignment results while SPs
    verify identified candidate locations and generate base-pair level
    alignments. Gray boxes indicate the intervals verified in the
    second round of verification (assuming $slice\_num = 4$).}
  \label{SPParallel}
\end{figure}

@@

\begin{figure}[!htb]
  \includegraphics[width=0.9\linewidth]{figures/MixPack}
  \caption{Bit-level encoding strategies for DNA sequences. For
    simplicity, we use a 4-bit word instead of 64-bit word for
    illustrating our combined AoSoA strategy (bottom).}
  \label{MixPack}
\end{figure}

@@

\begin{figure}[!htb]
  \includegraphics[width=1\linewidth]{figures/AsyncTrans}
  \caption{Asynchronous data transfer to LDM.}
  \label{AsyncTrans}
\end{figure}

@@
